\section{Sequential Gaussian processes latent variable model}
\label{sec:seqgplvm}
\hl{It seems that we dont need a first order CMM necessarily, we can have higher order CMMs as well? for now I 
will keep it first order because that is what Hatt et al do.}

Sequential Gaussian processes latent variable model (SeqGPLVM) is probabilistic latent variable 
model. It is based on the assumption that the conditional joint assigned treatment distribution 
can be written as a as a conditional Markov model (CMM) conditioned on both observed covariates and
the confounder substitute. 
\begin{equation}
    p_{\boldsymbol{\gamma}}(\bar{a}_{iT} \mid \bar{\mathbf{x}}_{iT}, \mathbf{z}_i) = p_{\gamma_0}(a_{i1} \mid \mathbf{x}_{i1}, \mathbf{z}_i) \prod_{t=2}^T p_{\gamma_t}(a_{it} \mid a_{i,t-1}, \mathbf{x}_{it}, \mathbf{z}_i)
\end{equation}
where \(\gamma = \{\gamma_0, \gamma_2, \ldots, \gamma_T\}\) are the \emph{generative} 
parameters of the treatment assignment model at each time point. Note that the each time point can have 
its own set of parameters. This is cruicial to understand the SeqGPLVM architecture. 
In theorem 1 of \cite{hattSequentialDeconfoundingCausal2024}, it is shown that, if the treatment 
assignment follows a CMM (for rigorous definition of CMM we can look at \cite{bouchattaouiCausalDynamicVariational2025}), we obtain sequential ignorability given the observed covariates and the
confounder substitute, i.e., for all $t \in \{1, \ldots, T\}$ and for all $\bar{a}_t$,
\begin{equation}
    \label{eq:seqgplvm-single-unit-seqignorability}
    Y_{i,t+1}(\bar{a}_t) \perp A_t \mid \bar{X}_{it}, \bar{A}_{i,t-1} ,Z_{i}.
\end{equation}

In our setting with one outcome measurement at the end of the study, we only need this 
result for t=T, i.e.,
\begin{equation}
    \label{eq:seqgplvm-single-unit-seqignorability-t=T}
    Y_{i,T+1}(\bar{a}_T) \perp A_t \mid \bar{X}_{iT}, \bar{A}_{i,T-1} ,Z_{i}.
\end{equation}
Note that this equation means that the potential outcome of the entire treatment history at time T+1 is independent of the treatment assignment 
at \emph{any} time point t, given the entire observed covariate history and treatment history up to time T-1, along with the confounder substitute. 

In \cite{hattSequentialDeconfoundingCausal2024}, the exact procedure for adjusting for unmeasured time-invariant confounding using SeqGPLVM is as follows:
\begin{itemize}
    \item \textbf{Stage 1: Sequential Latent Factor Model.} Fit a latent variable model to the observed treatment assignments and covariates over time to estimate the confounder substitute $\mathbf{z}_i$ for each individual.
    \item \textbf{Stage 2: Outcome Model.} Use the estimated confounder substitutes from Stage 1 as additional covariates to adjust for confounding in the outcome model. This outcome model can be MSM, G-formula, etc.
\end{itemize}
The key idea here is that once we have estimated the confounder substitutes $\mathbf{z}_i$, we can 
treat them as \emph{observed-like} covariates and include them in the outcome model to adjust for 
confounding, leveraging the sequential ignorability result in equation 
\ref{eq:seqgplvm-single-unit-seqignorability-t=T}. Despite the flexibility of this approach in choosing 
the outcome model, it has critical pitfalls that we will discuss in section \ref{sec:hatt-problems}. 
To avoid these issues, in our implementation, we use the estimated propensities from the
SeqGPLVM treatment model directly in the MSM using IPTW. This choice is also conceptually closer to 
the FE-MSM method described in section \ref{sec:fe-msm}. 

The sequential latent variable model in stage 1 is implemented using Gaussian processes latent variable models (GPLVMs) \cite{lawrenceGaussianProcessLatent2003}. 
Specifically, we place GP priors on the functions that map the latent variables and 
observed covariates to the treatment assignments at each time point, i.e., \(f_t \sim \mathcal{GP}(0, k_t(\cdot, \cdot))\) 
for \(t = 1, \ldots, T\), where \(k_t\) is the kernel function at time \(t\).
The use of GPs allows for flexible modeling of complex, non-linear relationships between the latent confounders, observed covariates, and treatment assignments over time.
