\section{Assumptions}
\label{sec:assumptions}
In this work, we categorize the assumptions into two groups: those necessary for 
the identification of causal effects and those required for estimation. 
As estimation assumptions are method-specific and highly technical, 
we refer the reader to the original papers for further detail. 
In this section, we outline the common \emph{identification} assumptions 
behind both the FE--MSM and SeqGPLVM approaches. Both methods described in this section 
rely on a set of assumptions to ensure the identifiability
of the causal effects in the presence of unmeasured time-invariant confounding. 
We summarize these assumptions below:
% --- Common assumptions (C1, C2, ...) ---
% Assumes you already defined the assumpC environment (C-prefix) in the preamble.

\begin{assumpC}[Consistency]\label{ass:C-consistency}
The observed outcome for each individual under their observed treatment history equals
the potential outcome under that treatment history. Formally, for each individual $i$
and time $t$, if $\bar{A}_{i,t}=\bar{a}_{i,t}$, then $Y_{i,t}=Y_{i,t}(\bar{a}_{i,t})$. 
Moreover, $Y_{i,T+1} = Y_{i,T+1}(\underline{a}_{i,T-k})$ if 
$\bar{A}_{i,T}=\left(\overline{A}_{i,T-k-1},\underline{a}_{i,T-k}\right)$.
\end{assumpC}

\begin{assumpC}[Positivity]\label{ass:C-positivity}
\[
\Pr\!\left[ A_t = a_t \,\middle|\, 
\overline{A}_{t-1} = \overline{a}_{t-1},\,
\overline{X}_t = \overline{x}_t \right] > 0
\]

for all values $(\overline{a}_{t-1}, \overline{x}_t)$ with

\[
\Pr\!\left[ \overline{A}_{t-1} = \overline{a}_{t-1},\,
\overline{X}_t = \overline{x}_t \right] \neq 0
\]

in the population of interest, and for all $a_t \in \mathcal{A}_t$.
\end{assumpC}

\begin{assumpC}[Time-invariance of unmeasured confounding]\label{ass:C-timeinvariant-uc}
The unmeasured confounding is assumed to be
time-invariant. That is, $\boldsymbol{U}_i$ does not change over time and
consistently affects both treatments across all time points and the final outcome.
\end{assumpC}

\begin{assumpC}[No interference]\label{ass:C-no-interference}
The treatment assignment for one individual does not affect the potential outcomes of
other individuals.
\end{assumpC}

