\section{Second stage: Outcome model}
\label{sec:outcome-model}

This section describes the second stage of the proposed two-stage estimation 
procedures. In the first stage, we fit a treatment model and obtain estimated propensity 
scores for each unit and time point. In the second stage, 
we use these propensities to construct inverse-probability weights and 
estimate functionals of the truncated mean response surface through a 
marginal structural model (MSM) parameterization.

Following \cite{blackwellEffectPoliticalAdvertising2024}, we focus on 
\emph{truncated} treatment histories: we consider interventions on the 
most recent $K+1$ treatment window while treating earlier treatment history as 
baseline confounding. Let $T$ denote the last observed period and fix $K\ge 1$. Define the recent treatment window
\[
\underline{A}_{i,T-K} := (A_{i,T-K},A_{i,T-K+1},\dots,A_{i,T}),
\]
and let $\underline{a}_{T-K}$ denote a counterfactual regime over that window. The corresponding truncated potential outcome is
\[
Y_i(\underline{a}_{T-K}) \equiv Y_i\!\big(\bar{A}_{i,1:T-K-1},\,\underline{a}_{T-K}\big),
\]
i.e., we intervene on the last window and leave earlier history as realized.

\subsection{Common framework: MSM parameterization and IPTW identification}
MSMs are \emph{marginal} models for counterfactual mean outcomes: 
rather than modeling the observed conditional outcome distribution, 
they parameterize the mean of potential outcomes directly as a 
function of the treatment regime \cite{robinsMarginalStructuralModels2000}.
Concretely, we posit an MSM of the form
\begin{equation}
\label{eq:msm-generic}
\mathbb{E}\!\left[Y_i(\underline{a}_{T-K})\right] \;=\; m(\underline{a}_{T-K};\gamma),
\end{equation}
where $m(\cdot;\gamma)$ is a user-chosen family indexed by a finite-dimensional parameter vector $\gamma$.

\paragraph{IPTW estimand and pseudo-population intuition.}
Both the FE-MSM and SeqGPLVM methods use Inverse Probability of Treatment Weighting (IPTW) 
to construct a pseudo-population where the treatment history is independent of the measured and 
unmeasured time-invariant confounders. Under the assumption of sequential ignorability 
(either unit-specific or conditioned on a latent substitute), the expected value of truncated potential outcome 
$Y_i(\underline{a}_{T-K})$ is identified using the following weighted estimand:
$$\mathbb{E}[Y_i(\underline{a}_{T-K})] = \mathbb{E}\left[\frac{\mathbf{1}(\underline{A}_{i,T-K} = \underline{a}_{T-K})Y_i}{\prod_{t=T-K}^{T} \pi_{it}^{a_t}(1 - \pi_{it})^{1-a_t}}\right].$$
In practice, the unknown $\pi_{it}$ are replaced with first-stage estimates $\widehat{\pi}_{it}$. 


\paragraph{Stabilized weights.}
To mitigate the high variance associated with unstabilized IPTW, 
we use \emph{stabilized} weights \cite{robinsMarginalStructuralModels2000}.
We adopt the standard longitudinal extension, using a numerator model that 
conditions on treatment history but excludes time-varying covariates, $p(A_{i,t} \mid \bar{A}_{i,t-1})$, and a 
denominator model that accounts for confounding (via observed history and the 
first-stage confounding adjustment). Concretely, our stabilized weight takes the form
\[
\widehat{SW}_i
=
\prod_{t=T-k}^{T}
\frac{
\widehat{\pi}^{\,\text{num}}_{it}{}^{A_{it}}\left(1-\widehat{\pi}^{\,\text{num}}_{it}\right)^{1-A_{it}}
}{
\widehat{\pi}^{\,\text{den}}_{it}{}^{A_{it}}\left(1-\widehat{\pi}^{\,\text{den}}_{it}\right)^{1-A_{it}}
},
\]
where $\widehat{\pi}^{\,\text{den}}_{it}$ is the first-stage propensity estimate used for confounding adjustment, and
$\widehat{\pi}^{\,\text{num}}_{it}$ is estimated from a reduced model depending only on treatment history.

\paragraph{Positivity and practical truncation.}
Identification and stable estimation require overlap (positivity): 
the probability of each relevant treatment history must be bounded away from $0$ and $1$. 
Although, we assume at the population level positivity holds (\ref{ass:C-positivity}), 
it does not guarantee positivity in realized finite samples.
When overlap is limited, weights can become extreme and estimates unstable; 
in the longitudinal MSM literature, this is a practical limitation of IPTW. 
To prevent numerically unstable weights, we apply propensity truncation by clipping all estimated probabilities to
\[
\widetilde{\pi}_{it} := \min\{1-\varepsilon,\;\max(\widehat{\pi}_{it},\varepsilon)\},\qquad \varepsilon=10^{-6}.
\]
In addition, for units with no within-unit treatment variation 
(always-treated or never-treated), estimating propensity estimation with individual effects can degenerate. 
Following \cite{blackwellEffectPoliticalAdvertising2024}, we handle these ``monotone'' 
units by setting the corresponding denominator probabilities to $1-\varepsilon$ 
(always-treated) or $\varepsilon$ (never-treated), ensuring finite weights and a 
well-defined second-stage fit. 

\paragraph{Second-stage estimation and uncertainty.}
Given weights $\widehat{SW}_i$, the second-stage estimator of $\gamma$ is 
obtained by fitting the MSM in
\eqref{eq:msm-generic} using a weighted estimating equation 
(equivalently, a weighted regression).
For valid uncertainty quantification under IPTW, we report robust (sandwich)
standard errors. As discussed in \cite{robinsMarginalStructuralModels2000},
non-robust model-based standard errors from weighted regressions are not generally
guaranteed to provide correct coverage, whereas robust (sandwich) intervals are
appropriate in this setting. This works for FE-MSM, however for SeqGPLVM, 
we also need to account for the additional sampling uncertainty from the latent substitute 
posterior. We describe our approach to this 
in the next section.

\subsection{FE-MSM}
The fixed-effects MSM (FE-MSM) addresses time-invariant unmeasured confounding by incorporating unit-specific effects directly into the first-stage propensity model. The identifying condition is \emph{unit-specific sequential ignorability}: conditional on observed history and a time-constant unit effect $\alpha_i$, treatment is as-if randomized at each time point.

\paragraph{First stage (propensity estimation).}
We estimate the propensity parameters and eliminate the nuisance unit effects via conditional maximum likelihood, producing fitted probabilities $\widehat{\pi}_{it}$ and thus stabilized weights for the second stage. This approach is motivated by panel asymptotics where longer panels alleviate bias from fixed-effect estimation noise.

\paragraph{Second stage (MSM fit).}
Using the stabilized and truncated weights described above, we estimate the MSM parameter $\gamma$ by solving the corresponding weighted estimating equations and report robust (sandwich) uncertainty.

\subsection{SeqGPLVM}
In contrast to the point-estimate adjustment in FE-MSM, the SeqGPLVM approach represents time-invariant unmeasured confounding through a latent substitute $Z_i$ and yields a posterior distribution $q(Z)$ from the first-stage sequential treatment model.

\paragraph{Connection to substitute-confounder logic.}
The identification strategy mirrors the general 
``substitute confounder'' idea: infer a latent variable that captures 
shared (multi-cause) confounding structure, while ruling out remaining 
single-cause confounding through an explicit assumption. In the multiple-cause 
setting, Wang and Blei formalize this as ``no unobserved single-cause confounders,
'' together with SUTVA and overlap. In our longitudinal setting, the SeqGPLVM plays 
the role of the assignment model that induces a substitute $Z_i$ used to adjust the 
treatment process.

\paragraph{Avoiding circularity in propensity modeling.}
To avoid the circular dependency that would arise from using $Z_i$ as a covariate in a separate propensity regression, we use the treatment probabilities implied by the SeqGPLVM directly as the denominator propensities in the stabilized weights.

\paragraph{Propagating latent uncertainty (MI-style MSM fitting).}
To propagate uncertainty about the latent substitute into the causal effect estimates, we use a multiple-imputation (MI)-style procedure: draw $S$ samples $Z^{(s)}\sim q(Z)$, compute draw-specific propensities and weights, and fit the same MSM to obtain $\widehat{\gamma}^{(s)}$ with robust covariance $V^{(s)}$. We pool the results using Rubin's within/between decomposition, which adds the between-draw variability to the average within-draw variance to reflect uncertainty from the latent first stage.

\paragraph{Remark on overlap.}
As in any IPTW-based procedure, limited overlap can lead to high-variance estimates. 
In the deconfounder framework, overlap is an explicit requirement for identifying mean 
potential outcomes given the substitute confounder. 
analogously, we enforce practical truncation of estimated probabilities to 
stabilize estimation in finite samples.

\subsection{FE-MSM}
The Fixed-Effects MSM (FE-MSM) addresses time-invariant unmeasured confounding by incorporating unit-specific fixed effects directly into the propensity score model.

Unit-Specific Sequential Ignorability: This method relies on the assumption that treatment is randomized conditional on the observed history and a time-constant unit-specific effect $\alpha_i$.
Estimation Procedure: We obtain estimates for the propensity parameters $\beta$ and the unit effects $\alpha_i$ via **Conditional Maximum Likelihood**. This approach is justified under **large-$N$, large-$T$ asymptotics, where the bias from noisy fixed-effect estimation (the incidental parameters problem) fades over time.
Weighted Least Squares: Once the fixed-effect weights are constructed, the MSM parameters $\gamma$ are estimated by solving a weighted estimating equation, which reduces to **Weighted Least Squares (WLS)** for our linear specification. We use a **robust (sandwich) variance estimator** to obtain valid Wald confidence intervals.

\subsection{SeqGPLVM}

In contrast to the point-estimate approach of FE-MSM, the SeqGPLVM captures unmeasured confounding by inferring a posterior distribution $q(Z)$ over a latent substitute $Z_i$.

Handling Circular Dependency: To avoid the "circular dependency" created by using $Z_i$ as a covariate in a separate propensity model, we use the estimated propensities from the SeqGPLVM **directly** to build the IPTW weights.
Multiple Imputation (MI)-Style Procedure: To propagate the latent uncertainty from the GP-based treatment model, we draw $S$ samples $Z^{(s)} \sim q(Z)$. For each draw, we:
    1.  Compute draw-specific propensities and weights $W^{(s)}$.
    2.  Fit the weighted MSM to obtain a draw-specific estimate $\hat{\gamma}^{(s)}$ and robust covariance $V^{(s)}$.
Rubin’s Rules for Pooling: The final point estimate is the average across draws, $\bar{\gamma} = S^{-1}\sum \hat{\gamma}^{(s)}$. The total variance is calculated using **Rubin’s within/between decomposition**, $T = W + (1 + 1/S)B$, which explicitly accounts for the variability induced by the uncertainty in the inferred latent confounder.



Although the seqgplvm can be used in the settings in which the outcome is observed at every time point, 
since the asymptotic properties and consistency of their IPTW-FE estimator 
relies on having less number of N than T (assumption )
Based on the consistency assumption \ref{ass:C-consistency}, we can define 
truncated potential outcomes 