\chapter{Discussion}
\label{discussion}
\begin{itemize}
    \item discuss about the vocablurary choice of "unmeasured" vs "hidden"
    \item testing against misspecified latent space dimension (I think bica does this?)
    \item implications of having one GP per time step vs one big GP for all time steps (maybe cite 
    some GP time series work?)
    \item discuss the none parametric nature of GP and its implications, specifically in the context of IPTW and the assumption of correct propensity model specification
    \item discuss about the amortized version of GPLVM (perhaps in the appendix)
    \item \hl{what} \cite{hattSequentialDeconfoundingCausal2024} \hl{did seems very wrong. it seems that they extract the latent variables from the treatment model and then use them as if they were observed covariates in the MSM. The worrisome part is that they are using IPTW which means after extracting the esimated latent substitute, they fit another propensity model incluing the estimated treatments so basically on the left hand side we have $a_it$ and the on the right hand side we have a function of $a_it$. This seems very wrong. We should discuss this. What I did, which seems more correct, is to use the estimated propensities from the seqgplvm treatment model directly in the MSM. Althoughm, this can also cause some problems because seqgplvm is a very powerful model and it can overfit the treatment assignment and lead to extreme weights.}
    \item another break point of this whole story of deconfounding with latent variables is the assumption of pinpointability. in the context of the time varying treatments, this assumption means that the latent variables estimated should stay the same for all hypothetical treatment assignments. I need to read D'amor critique of deconfounding paper again. 
    \item another big issue is that, their second propensity model is the msm stage is not even correctly specified! Although they removed the explicit equations they used for the MSM stage, from their previous version, I remeber that they used a model that was not the same as the DGP. 
    \item Another point is that, the selling point of Hatt et al was that SeqGPLVM is flexible in terms of 
    captuing non linear relationships between the latent confounders, observed covariates and treatment assignments. However, 
    because of the pinpointability assumption, there is a high chance that the those more complicated relationships in the dgp, 
    cannot satisfied the pinpointability assumption. It would be nice if I can find a dgps that is not very complicated but 
    violates the pinpointability assumption. Maybe reading D'amor critique paper again can help with this.
\end{itemize}

\section{Problem of reusing the estimated latent variables in a second stage outcome model involving propensity score estimation}
\label{sec:hatt-problems}
As mentioned in section \ref{sec:seqgplvm}, in \cite{hattSequentialDeconfoundingCausal2024}, the authors treat 
SeqGPLVM as a factor model whose purpose is to estimate the posterior over the latent confounder substitutes, 
\(p(\mathbf{z}_i \mid \bar{a}_{iT}, \bar{\mathbf{x}}_{iT})\). Then, the authors proceed to use samples 
from this posterior distribution as if they were observed covariates in a second stage outcome model. 
Specifically, they suggest fitting two two propensity score based models: 
marginal structural model (MSM)\cite{robinsEstimationCausalEffects2008} and 
Recurrent Marginal Structural Networks \cite{limForecastingTreatmentResponses2018}. The 
problem with this approach is that in both of these models, the authors fit a second propensity score model 
that includes the estimated confounder substitutes \(\mathbf{z}_i\) as covariates. However, since \(\mathbf{z}_i\) 
are estimated from the treatment assignments \(\bar{a}_{iT}\), including them as covariates in a second propensity score 
model that aims to predict \(\bar{a}_{iT}\) creates a circular dependency. More concretely, 
in the second stage propensity score model, we are trying to estimate 
\(p(A_{it} \mid \bar{A}_{i,t-1}, \bar{\mathbf{X}}_{it}, \mathbf{z}_i)\) where \(\mathbf{z}_i\) is a function of 
\(\bar{a}_{iT}\) and \(\bar{x}_{iT}\). 
This circular dependency can lead to biased estimates of
using inverse probability of treatment weighting (IPTW), where the propensity scores are estimated by including the estimated confounder substitutes $\mathbf{z}_i$ as covariates.

Also note that the correct way of using the posterior distribution over 
the latent confounder substitutes \(p(\mathbf{z}_i \mid \bar{a}_{iT}, \bar{\mathbf{x}}_{iT})\)
is to integrate over this distribution when estimating the causal effect, i.e.,
\begin{equation}
    \mathbb{E}\left[Y_{iT}(\bar{d}_{T-k:T})\right] = \int \mathbb{E}\left[Y_{iT}(\bar{d}_{T-k:T}) \mid \mathbf{z}_i\right] p(\mathbf{z}_i \mid \bar{a}_{iT}, \bar{\mathbf{x}}_{iT}) d\mathbf{z}_i
\end{equation}
which can be approximated using Monte Carlo integration by drawing samples from the posterior distribution, 
calculate the conditional propensity scores by the SeqGPLVM likelihood for the 
observed treatment, and then use these propensity scores in the IPTW estimation of the MSM.
Repeat this for multiple samples from the posterior and average the results to obtain the final estimate of the causal effect. 

Note that in \cite{bicaTimeSeriesDeconfounder2020} and \cite{wangBlessingsMultipleCauses2019} 
They also follow a similar two stage procedure where they first estimate the latent confounders
using a factor model and then use these estimated confounders as observed covariates in the outcome model. 
However, they do not use IPTW for estimating the causal effect in the outcome model. 
Instead, they use G-formula which does not involve fitting a second propensity score model.

