\chapter{Notation and setup}
In this section we introduce the notation and the setupused throughout this thesis. 
Let $i=1, \ldots, N$ index the individuals in the study and let $t=1, \ldots, T$ 
index the time points at which measurements are taken. We denote the treatment 
variable at time $t$ for individual $i$ as $A_{it}$, which can take on values from a 
predefined set of treatment options. In this thesis, 
we assume a binary treatment, $A_{it} \in \{0,1\}$. We adopt the convention that treatment 
is absent at the beginning of the study, $t=0$ and therefore $A_{i0} = 0$ 
for all individuals $i$. 
The outcome variable at 
time $t$ for individual $i$ is 
denoted as $Y_{i,t+1}$ to emphasize that the outcome is measured after the treatment at time $t$.
We also consider a set of observed covariates $X_{it}$, 
which may include both time-varying 
and time-invariant variables. The history of treatment and covariates up to time $t$ for 
individual $i$ is represented as
$\bar{A}_{it} = (A_{i1}, A_{i2}, \ldots, A_{it})$ and $\bar{X}_{it} = (X_{i1}, X_{i2}, \ldots, X_{it})$, 
respectively. We denote the potential outcome at time $t$ for individual $i$ under a specific treatment
regime $\bar{a}_t = (a_1, a_2, \ldots, a_t)$ as $Y_{i,t+1}(\bar{a}_t)$. Finally, we use $U_i$ 
to represent \emph{structural} unmeasured confounders that are time-invariant for individual $i$. 

We also define truncated versions of the treatment and covariate histories from \(k\) 
steps prior up to time \(t\) as \(\underbar{A}_{i,t-k}\) and \(\underbar{X}_{i,t-k}\), respectively. 
The outcome setup we are focusing on is a single outcome measured at the end of 
the study, i.e., at time T+1, denoted as $Y_{i,T+1}$. This is the setup considered in
\cite{blackwellEffectPoliticalAdvertising2024} and it is a special case of the setup in
\cite{hattSequentialDeconfoundingCausal2024} who consider outcomes measured at each time point. This 
choice is both motivated by the specific application they consider 
(political advertising) and technical reasons related to the consistency of the IPTW-FE estimator.

Moreover, since the asymptotic properties and consistency of the IPTW-FE estimator 
proposed by \cite{blackwellEffectPoliticalAdvertising2024} relies on the 
truncated potnetial outcomes, we can define the truncated potential outcomes as
$Y_{i,T+1}(\underline{a}_{T-k}) \equiv Y_{i,T+1}(\overline{A}_{i,T-k-1},\, \underline{a}_{i,T-k})$. This 
definition is to capture the meaning of the potential outcome of manipulating 
the last $k$ treatment values, while leaving the earlier treatment values as observed. In this way, 
the earlier treatments before the last $k$ time points can be thought of as part of the baseline covariates. 