\chapter{Simulation study}
\label{sec:simulation_study} 
\section{Data generating process}
In this section, we present a simulation study designed to evaluate the performance 
of the two methods for estimating causal effects with time-varying treatments 
under time-invariant unmeasured confounding. The data generation process (DGP) is taken 
from the DGP described in Section 5.1 of \cite{blackwellEffectPoliticalAdvertising2024}. In this 
DGP, the scenario which we simulate is a study design where individuals are exposed to a binary treatment
(e.g., receiving a political advertisement) over a period fixed period of time, and the outcome of interest
is measured only once at the end of the study period (e.g., voting behavior). The treatment assignment
is influenced by both observed covariates and an unmeasured time-invariant confounder (e.g., political
leaning). The outcome is affected by the treatment received at the final time point, the cumulative 
effect of treatments received in the last three time points, the mean of observed covariates over time, 
and the unmeasured confounder. 

We simulate longitudinal data for $n \in \{200, 500, 1000, 3000\}$ individuals. For the
length of the time serieses for each n, we consider three different lenghts: $\rho = n/T$ 
with $\rho \in \{5,10,50\}$. The data is balanced. menaning that each individual has 
the same number of time points. In Eq.~\eqref{eq:blackwell_scm}, we present 
the structural causal equations for the DGP. Here, $\operatorname{expit}(x) = \frac{1}{1 + e^{-x}}$ and 
$\bar{X}_{iT} = \frac{1}{T}\sum_{t=1}^T X_{it}$.

\begin{subequations}\label{eq:blackwell_scm}
\begin{align}
U_i &:= a(2V_i-1), \qquad V_i \sim \operatorname{Unif}(0,1)
\\
X_{it} &:= -\tfrac12\mathbbm{1} + L \xi_{it} \qquad \xi_{it}\sim \mathcal{N}(0,I), \qquad LL^\top=\Sigma
\\
A_{it} &:= \mathbbm{1}\!\left\{\eta_{it} \le
\operatorname{expit}\!\left(U_i+\varphi A_{i,t-1}+\beta^\top X_{it}\right)\right\},
\qquad \eta_{it} \sim \operatorname{Unif}(0,1)
\\
Y_{iT} &:= U_i + \tau_F A_{iT} + \tau_C \sum_{t=T-3}^{T-1} A_{it}
+ \gamma^\top \bar X_{iT} + \epsilon_{iT},
\qquad \epsilon_{iT} \sim \mathcal{N}(0,1)
\label{eq:blackwell_scm_outcome}
\end{align}
\end{subequations}

We simulate for two different values of $a \in \{1, 2\}$ \hl{to control the
contribution of the unmeasured confounder $U_i$ to the variance of the linear predictor.}. For the number 
of covariates, we simulate for $p \in \{2, 4\}$. 
The coefficients are set to $\beta = (-0.5, -0.5)^{\top}$ when $p=2$ and 
$\beta = (-0.5, -0.5, 1, -0.5)^{\top}$ when $p=4$. \hl{(add some note about why we chose 
these coeffs based on the variance or sth)}. We set $\varphi = 0.3$. Note that we generate 
the covariates $X_{it}$ exogenously to the treatments. So we do not have the 
treatment-confounder feedback. This is done to isolate the effect of
adjusting for unmeasred time-invariant confounding and make the simulation simpler and it has important 
implications in interpreting the causal effects. see section \ref{estimation}. Also note that theere is no dynamics 
in the covariates and the distribution of $X_{it}$ is the same for all time points. The parameters 
for the outcome model are set to $\tau_F = 1$, $\tau_C = 0.3$, and $\gamma = (1, 0.5)^{\top}$ 
when $p=2$ and when $p=4$, $\gamma = (1, 0.5, 1, 1)^{\top}$.

\section{Estimation}
\label{estimation}
For defining our estimation goals, first we need specify the causal question of interest. The causal
question we want to answer is: What is the average effect of the treatment received at the final time on the 
final outcome, fixing the last $K$ treatments before $T$: 
$\mathbb{E}\left[ Y_{iT}\left(d_T = 1, \bar{d}_{i,T-K:T-1}\right) - Y_{iT}\left(d_T = 0, \bar{d}_{i,T-K:T-1}\right) \right]$, and the 
average effect of receiving one additional treatment in the last three time points, before $T$ on the final outcome, fixing 
the last treatments. To express 
this mathematically, we first define $S_k(\bar{d}_{T-K:T-1}) := \sum_{t=T-K}^{T-1} d_t \in \{0,1,2,\dots, K\}.$ as the sum 
of treatments received in the last $K$ time points before $T$. Then, the average effect of receiving one additional treatment 
in these $K$ time points on the final outcome is defined as 
$\mathbb{E}\left[Y_{iT}\left({S_k(\bar{d}_{T-K:T-1}) = s, d_{iT}}\right) - Y_{iT}\left({S_k(\bar{d}_{T-K:T-1}) = s+1, d_{iT}}\right)\right], s \in \{0,1,\dots,K-1\}$. 
To answer both of these causal questions, we can propose a marginal structural model (MSM) 
that relates the treatment history to the expected potential outcome at the final time: 
$\mathbb{E\left[Y_{iT}\left(\bar{d}_{T-k:T}\right)\right]} = g(\bar{d}_{T-k:T}, \gamma_0)$, where 
$g(\bar{d}_{T-k:T}, \gamma_0)$ is the true unkown function parameterized by $\gamma_0$. Our job is 
to first propose a function $f(\bar{d}_{T-k:T}, \gamma)$ and then estimate its parameters $\gamma$. In this 
DGP, the true function $g$, is: 
\begin{equation}
    \label{truemsm}
    g\left(\underline{d}_{T-k}; \gamma_0 \right) = \alpha_0 + \tau_{F} d_T + \tau_{C} \sum_{t=T-3}^{T-1} d_t
\end{equation}

to see this, we begin from Eq.~\eqref{eq:blackwell_scm_outcome}. First note that since this is 
a sctructural equation model, we can replace the random variables by their interventional values. So we have:
\begin{align}
\label{truemsmproof}
Y_{iT}\!\left(\bar d_{T-k:T}\right)
&= U_i + \tau_F d_T + \tau_C \sum_{t=T-3}^{T-1} d_t + \gamma^\top \bar X_{iT} + \epsilon_{iT} \\
\implies\quad
\mathbb{E}\!\left[Y_{iT}\!\left(\bar d_{T-k:T}\right)\right]
&= \mathbb{E}[U_i] + \tau_F d_T + \tau_C \sum_{t=T-3}^{T-1} d_t
   + \gamma^\top \mathbb{E}[\bar X_{iT}] + \mathbb{E}[\epsilon_{iT}] \\
\implies\quad
\mathbb{E}\!\left[Y_{iT}\!\left(\bar d_{T-k:T}\right)\right]
&= \mathbb{E}\!\left[U_i + \gamma^\top \bar X_{iT}\right]
   + \tau_F d_T + \tau_C \sum_{t=T-3}^{T-1} d_t \\
\implies\quad
\mathbb{E}\!\left[Y_{iT}\!\left(\bar d_{T-k:T}\right)\right]
&= \alpha_0 + \tau_F d_T + \tau_C \sum_{t=T-3}^{T-1} d_t,
\qquad
\alpha_0 := \mathbb{E}\!\left[U_i + \gamma^\top \bar X_{iT}\right].
\end{align}

\hl{The important point to note here is that since $X_{it}$ is exogenous to the treatment, we do not have 
terms involving $X_{it}(\underline{d}_{T-k})$ which are counterfactual covariates under 
the specified treatment regime and we could substitute it with a constant.} From 
Eq.~\eqref{truemsmproof}, we can see 
$\mathbb{E}\left[ Y_{iT}\left(d_T = 1, \bar{d}_{i,T-K:T-1}\right) - Y_{iT}\left(d_T = 0, \bar{d}_{i,T-K:T-1}\right) \right] = \tau_F$ and 
$\mathbb{E}\left[Y_{iT}\left({S_k(\bar{d}_{T-K:T-1}) = s, d_{iT}}\right) - Y_{iT}\left({S_k(\bar{d}_{T-K:T-1}) = s+1, d_{iT}}\right)\right] = \tau_C$. 
The equality between these causal effects and the parameters of the MSM is a result of the linearity of the DGP and the absence of treatment-confounder feedback. 
See \ref{sec:msm_ground_truth} for more discussion on this.

To only focus on the adjustment for the unmeasred confounder, for both of the methods, 
we specify $f\left(\underline{d}_{T-k}; \gamma \right)$ as the correctly specified linear MSM:
\begin{equation}
    f\left(\underline{d}_{T-k}; \gamma \right) = \alpha' + \tau_F' d_T + \tau_C' \sum_{t=T-3}^{T-1} d_t
\end{equation}
where $\gamma = (\alpha', \tau_F', \tau_C')$. 
\subsection{Fix-effect marginal structural model}

\section{Results}



 
 



