---
title: "R Notebook"
output: html_notebook
---

```{r}
library(lme4)
library(dplyr)
library(jsonlite)
library(arrow)
library(dplyr)
library(ggplot2)

train_id  <- "281b276cad"
train_out <- file.path(getwd(), "results", "models", "seqgplvm", train_id)

# 1) read data_ref.json (lives in train_out)
path <- file.path(train_out, "data_ref.json")
j <- read_json(path, simplifyVector = TRUE)

# 2) load the parquet (data_file is a DIR that contains data.parquet)
df <- arrow::read_parquet(file.path(getwd(), j[["data_file"]], "data.parquet"))
df_manifest <- read_json(file.path(getwd(), j[["data_file"]], "manifest.json"), header = TRUE,  sep = '	',  stringsAsFactors = FALSE)
df_manifest$params
# 3) read the split file and extract ids
splits <- read_json(file.path(getwd(), j[["split_file"]]), simplifyVector = TRUE)
train_split <- splits[["train_ids"]]
val_split   <- c(splits[["val_ids"]], splits[["test_ids"]])
# 4) filter df to only training ids
df_train <- df %>% filter(patient_id %in% train_split)
# Fit textbook GLMM: logit(p_it) = α_id + φ * Dlag + β1*x1 + β2*x2
# GLMM (with random intercepts)
fit_re <- glmer(D ~ lag1_D + x0 + x1 + (1 | patient_id),
                data = df_train, family = binomial(link = "logit"))

# GLM (no random effects)
fit <- glm(D ~ lag1_D + x0 + x1,
              data = df_train, family = binomial(link = "logit"))

summary(fit_re)        # coefficients + random-effect SD
summary(fit)        # coefficients only
fixef(fit_re)          # (Intercept), Dlag, x1, x2

re <- ranef(fit_re)$patient_id
alpha_map <- setNames(re[,"(Intercept)"], rownames(re))
df_train$alpha_hat <- alpha_map[ as.character(df_train$patient_id) ]       # α_i estimates (BLUPs)
df_train$phat <- predict(fit, type = "response")
df_train$phat_re <- predict(fit_re, type = "response")

corr = df_train %>%
  distinct(patient_id, alpha, alpha_hat) %>%
  summarise(corr = cor(alpha, alpha_hat))


df_train %>%
  distinct(patient_id, alpha, alpha_hat) %>%
  ggplot(aes(alpha, alpha_hat)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(x = "True αᵢ", y = "Estimated α̂ᵢ",
       title =sprintf("Correlation = %.3f", corr)
    )
sigma_re <- attr(VarCorr(fit)$id, "stddev")  # SD of random intercepts

# Predicted probabilities for each row:
corr
df_plot <- df_train |> dplyr::mutate(phat = phat)
#df_plot <- df_train |> dplyr::mutate(phat = phat_re)

ggplot(df_plot, aes(x = p_true, y = phat)) +
  geom_abline(slope = 1, intercept = 0, linetype = 2) +
  geom_point(alpha = 0.3, size = 1) +
  coord_equal(xlim = c(0,1), ylim = c(0,1)) +
  labs(x = "True p", y = "Predicted p̂", title = "True vs Predicted Probabilities")
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
```{r}

```

